---
title: "myPay Analysis"
author: "Paul Testa"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r init,echo=F}
library(knitr)
opts_chunk$set(tidy=TRUE,echo=TRUE,results='markup',strip.white=TRUE,cache=FALSE,highlight=TRUE,width.cutoff=132,size='footnotesize',out.width='1.2\\textwidth',message=FALSE,warning=TRUE,comment=NA)

options(width=110,digits=4)
```



# Agency Objective

Increase military retirees’ reactivation of their online military retirement finance management platform accounts to allow retirees faster access to their tax documents and to offset operational burden for the agency during tax season.

# Background

One of the purposes of the online platform myPay is to help military retirees manage their military retirement finances. It allows electronic download of tax documents, such as a Retirement Account Summary and the IRS-1099 form. While myPay accounts are automatically created for all military retirees, many have never used their account, or have let their accounts deactivate by not logging in for 180 consecutive days. At the time of this pilot, about 55% of the retiree population had inactive myPay accounts (1.4 out of 2.5 million retirees). This poses a problem during tax season, when the number of retirees contacting the Defense Finance and Accounting Service (DFAS) surges as retirees try to either re-gain access to myPay or request copies of their tax documents. For example, in the fourth quarter of 2014 DFAS received a monthly average of 12,393 calls regarding accessing myPay accounts, compared to 29,046 during the first quarter (tax season) of 2015. The increased call volume during tax season places operational burden on DFAS and may result in slower services for military retirees. 

# Design

DFAS and SBST designed nine different emails to encourage reactivation of myPay accounts prior to the beginning of tax season. A baseline email template was created using existing language about myPay from DFAS’ newsletters and letters to retirees, with added clear action steps on how to reactivate a myPay account.^[Clear information can induce action. Hastings, J., & Weinstein, J. (2008). Information, School Choice, and Academic Achievement: Evidence from Two Experiments. The Quarterly Journal of Economics, 123(4), 1373-1414. Also see TSP SmartDocs pilot in SBST 2015 Annual Report for another example of clear action steps.]  Eight additional email variants were built upon the baseline template using behavioral insights: an email with a DFAS team signature versus an email with a DFAS director’s signature, one email highlighting the resources that could be given to current Servicemembers by cutting DFAS’ mailing costs, one email appealing to the informational security of using myPay, one email that provided social proof information,^[“Social proof” refers to the use of information on the common use of myPay by a large number of military retirees, from which individuals who are considering using myPay might infer its usefulness. ]  and three emails with different framings of how myPay can simplify retrieval of tax documents.^[A generic statement about access to “tax documents,” a statement instead referring to specific tax documents – the “1099 and RAS” – and a statement asking individuals to think about their experience last tax season.]  A control group did not receive any email during testing, but received an email after the end of the pilot.^[After the pilot, two emails were chosen from the original nine, and sent to the control group.]^[Including the control group, 261,754 retirees participated in the pilot.]

Emails were sent out in four separate waves between September 8 and November 5, 2015, two to three treatments at a time.^[Emails were mostly sent on Tuesdays, with the exception of the final two emails, which went out on Wednesday November 4th and Thursday November 5th.]

**Table 1. Treatment Conditions in Retiree myPay Account Reactivation Pilot**

| Treatment Group | Treatment                       | Date Sent | N     |
|---------------- |:-------------------------------:|-----------|------:|
| Group 1         | Control: No initial email$^a$       | 8-Dec-15$^a$  | 49496 |
| Group 2         | Baseline Template               | 22-Sep-15 | 23637 |
| Group 3         | Signature: Individual           | 4-Nov-15  | 23593 |
| Group 4         | Signature: Group                | 5-Nov-15  | 23538 |
| Group 5         | Tax Time Benefits: General      | 22-Sep-15 | 23551 |
| Group 6         | Tax Time Benefits: Specific     | 22-Sep-15 | 23765 |
| Group 7         | Tax Time Benefits: Past Hassles | 9-Sep-15  | 23580 |
| Group 8         | Benefits to Troops              | 20-Oct-15 | 23700 |
| Group 9         | Increase in Security            | 20-Oct-15 | 23455 |
| Group 10        | Social Proof                    | 20-Oct-15 | 23439 |

$^a$ After the initial treatment emails were sent, subject in the control condition were randomly assigned to receive the treatment of Group 6 (Tax Time Benefits: Specific), Group 9 (Increase in Security) or a combination of the messages in Group 6 and 9.

Subjects in a  list of 212,258 email addresses were completely randomly assigned to 1 of 10 conditions (9 treatment and 1 control).  Emails were sent  in waves.   Each wave contains two to three treatment conditions. Specifically:

**Table 2. Treatment Assignment by Waves**

| Wave | Date             | Treatments        | N     |
|------:|------------------:|-------------------:|-------:|
| 1    | 9/9/15           | Group 7           | 23580 |
| 2    | 9/22/15          | Groups 2, 5, & 6  | 70953 |
| 3    | 10/20/15         | Groups 8, 9, & 10 | 70594 |
| 4    | 11/4/15; 11/5/15 | Groups 3 & 4 | 47131 |

# Analysis Plan

Here is the plan for the analysis. We agreed on this plan before doing analysis.


## Data and Outcomes

Six to nine days after each treatment wave was sent, DFAS queried their myPay records  providing an updated list of who had accessed and reactivated their myPay accounts.  The outcome for a particular treatment is defined from these multiple datasets as the proportion of recipients who accessed their myPay account over a six-day period starting on the day a given email treatment was sent. This outcome can be compared to  the proportion of participants accessing their myPay accounts  over the same number of days (but at a potentially different point in time)  in the following groups:

- All untreated participants (Participants who have not yet received an email)
- All “control” participants (Group 1)  who did not receive an email until December 8, 2015.
- Subjects in the baseline treatment (Group 2)
- Subjects in a similar treatment condition  (e.g. Group 3 vs Group 4)

Table 4 details the dates data were collected, the treatments administered at that time  and  the relevant comparisons for each treatment. Note the date ranges for some comparisons involve comparisons across different waves of the study.

**Table 4: Dates of treatment administration and outcome observation**

+-------+-------------------+---------------+---------------------------+
| Wave 	| Date            	| Treatment     | Relevant Comparison      	|
+=======+===================+===============+===========================+
| 1	    |**Administered:**	| Group 7		| **All Untreated**		  	|
|	    |			  		|				|							|
| 	    |- 9/9/15 	   		| 				| -	Groups 1-6, 8-10 	  	|
|	    |			   		|				| - Dates: 9/9-9/14			|
|	    |**Observed:**   	| 				| - Wave: 1					|
|	    |			   		|				|							|
|	    |- 9/18/15	   		|				| **Control**				|
|	    |			   		|				|							|
| 	    |**Outcome:**   	| 				| -	Groups 1		 	  	|
|	    |			   		|				| - Dates: 9/9-9/14			|
|	    |- 9/9-9/14	  		| 				| - Wave: 1					|
|	    |			   		|				|							|
|	    |			   		|				| **Baseline**				|
|	    |			   		|				|							|
| 	    |		  	  		| 				| -	Groups 2		 	  	|
|	    |			   		|				| - Dates: 9/22-9/27		|
|	    |		 	   		| 				| - Wave: 2					|
|	    |			   		|				|							|
|	    |			   		|				| **Similar**				|
|	    |			   		|				|							|
| 	    |		  	   		| 				| -	Groups 5,6		 	  	|
|	    |			   		|				| - Dates: 9/22-9/27		|
|	    |		 			| 				| - Wave: 2					|
+-------+-------------------+---------------+---------------------------+
| 2	   	|**Administered:**	| Group 2,5,&6	| **All Untreated**		  	|
|	  	|			   		|				|							|
| 	  	|- 9/22/15 	   		| 				| -	Groups 1,3,4,8-10 	  	|
|	  	|			   		|				| - Dates: 9/22-9/27		|
|	  	|**Observed:**  	| 				| - Wave: 2					|
|	  	|			   		|				|							|
|	  	|- 9/28/15	   		|				| **Control**				|
|	  	|			   		|				|							|
| 	  	|**Outcome:**    	| 				| -	Groups 1		 	  	|
|	  	|			   		|				| - Dates: 9/22-9/27		|
|	  	|- 9/22-9/27	   	| 				| - Wave: 2					|
|	  	|			   		|				|							|
|	  	|			   		|				| **Baseline**				|
|	  	|			   		|				|							|
| 	  	|		  	 		| 				| -	Groups 2		 	  	|
|	  	|			   		|				| - Dates: 9/22-9/27		|
|	  	|		 			| 				| - Wave: 2					|
|	  	|			   		|				|							|
|	  	|			   		|				| **Similar**				|
|	  	|			   		|				|							|
| 	  	|		  			| 				| -	Groups 5; 6,7	 	  	|
|	  	|			   		|				| - Dates: 9/9-9/14; 9/22-9/27|
|	  	|		 			| 				| - Wave: 1; 2				|
+-------+-------------------+---------------+---------------------------+
| 3	   	|**Administered:**	| Group 8,9,&10	| **All Untreated**		  	|
|	   	|			   		|				|							|
| 	   	|- 10/20/15 	   	| 				| -	Groups 1,3,4	 	  	|
|	   	|			   		|				| - Dates: 10/20-10/25		|
|	   	|**Observed:**    	| 				| - Wave: 3					|
|	   	|			   		|				|							|
|	   	|- 10/28/15	   		|				| **Control**				|
|	   	|			   		|				|							|
| 	   	|**Outcome:**     	| 				| -	Groups 1		 	  	|
|	   	|			   		|				| - Dates: 10/20-10/25		|
|	   	|- 10/20-10/25  	| 				| - Wave: 3					|
|	   	|			   		|				|							|
|	   	|			   		|				| **Baseline**				|
|	   	|			   		|				|							|
| 	   	|		  	   		| 				| -	Groups 2		 	  	|
|	   	|			   		|				| - Dates: 9/22-9/27		|
|	   	|		 	   		| 				| - Wave: 2					|
|	   	|			   		|				|							|
|	   	|			   		|				| **Similar**				|
|	   	|			   		|				|							|
| 	   	|		  	   		| 				| -	Groups 8-10	 	  		|
|	   	|			   		|				| - Dates: 10/20-10/25		|
|	   	|		 	  		| 				| - Wave: 3					|
+-------+-------------------+---------------+---------------------------+
| 4	   	|**Administered:**	| Group 3;4 	| **All Untreated**		  	|
|	   	|			   		|				|							|
| 	   	|- 11/4;11/5    	| 				| -	Groups 1		 	  	|
|	   	|			   		|				| - Dates: 11/4-11/9; 11/5-11/10		|
|	   	|**Observed:**  	| 				| - Wave: 4					|
|	   	|			   		|				|							|
|	   	|- 11/10/15	   		|				| **Control**				|
|	   	|			   		|				|							|
| 	   	|**Outcome:**     	| 				| -	Groups 1		 	  	|
|	   	|			   		|				| - Dates: 11/4-11/9; 11/5-11/10		|
|	   	|- 11/4-11/9	 	| 				| - Wave: 4					|
|	   	|- 11/5-11/10	   	|				|							|
|	   	|			   		|				| **Baseline**				|
|	   	|			   		|				|							|
| 	   	|		  	  		| 				| -	Groups 2		 	  	|
|	   	|			   		|				| - Dates: 9/22-9/27		|
|	   	|		 	   		| 				| - Wave: 2					|
|	   	|			   		|				|							|
|	   	|			   		|				| **Similar**				|
|	   	|			   		|				|							|
| 	   	|		  	   		| 				| -	Groups 3,4		  		|
|	   	|			   		|				| - Dates: 11/4-11/9; 11/5-11/10		|
|	   	|		 	   		| 				| - Wave: 4					|
+-------+-------------------+---------------+---------------------------+



# Results

```{r}
data<-read.csv("data/mypay_clean.csv")
# Relevel treatment factors
data$treatment<-factor(data$treatment,levels=c("Control","2: Baseline",3:10))

```


Between September 8 and November 10, 2015, 5.5% of retirees that did not receive any emails reactivated their myPay accounts.

```{r}
with(data,apply(table(accessed,treat_anyemail),2,prop.table))
```


In comparison, 10.2% of retirees who received an email had reactivated their myPay accounts – a difference of 4.7 percentage points from the control group (p < 0.001, 95% CI [4.50,4.98]). 

```{r}
library(sandwich)
library(lmtest)
source("confintHC.R") 
m1<-lm(accessed~treat_anyemail,data)
coeftest(m1,vcov=vcovHC(m1,type="HC2"))
confint.HC(m1,level=.95,parm="treat_anyemailEmail",thevcov=vcovHC(m1,type="HC2"))
with(data,prop.test(table(treat_anyemail,accessed)[2:1,2:1]))
# Slightly different confidence intervals, same result

```

This translates to about 800-1,000 additional customer service calls during the 3.5 months of pilot. 

The most effective email appealed to the security of using myPay to retrieve tax documents, compared to using traditional mail.^[“Don’t risk getting your W-2 and other tax documents lost in the mail or opened by someone else.”]  This email increased reactivation by 5.5 percentage points compared to no email (p < 0.001, 95% CI [5.13,5.77]). 

```{r}
# Compare treatment effects over total study
with(data,apply(table(accessed,treatment),2,prop.table))
with(data,apply(table(accessed,treatment),2,prop.table))[,9]
m2<-lm(accessed~treatment,data)
coeftest(m2,vcov=vcovHC(m2,type="HC2"))["treatment9",]
confint.HC(m2,level=.95,parm="treatment9",thevcov=vcovHC(m2,type="HC2"))
table(data$treatment)
# Look specifically at Treatment 9 (increased security) vs Control
# During 6-day window in Wave 3 
data$treat9vC<-NA
data$treat9vC[data$treatmen=="Control"]<-"Control"
data$treat9vC[data$treatmen=="9"]<-"Treatment 9"
with(data, prop.test(table(treat9vC,accessed_w3)[1:2,2:1]))

m9vC<-lm(accessed_w3_c~treat9vC,data)
coeftest(m9vC,vcov=vcovHC(m9vC,type="HC2"))["treat9vCTreatment 9",]
confint.HC(m9vC,level=.95,parm="treat9vCTreatment 9",thevcov=vcovHC(m9vC,type="HC2"))

```




```{r}
# Other comparisons to pure control

# Wave 1
data$treat7vC<-NA
data$treat7vC[data$treatment=="Control"]<-"Control"
data$treat7vC[data$treatment=="7"]<-"Treatment 7"
m7vC<-lm(accessed_w1_c~treat7vC,data)
coeftest(m7vC,vcov=vcovHC(m7vC,type="HC2"))["treat7vCTreatment 7",]
confint.HC(m7vC,level=.95,parm="treat7vCTreatment 7",thevcov=vcovHC(m7vC,type="HC2"))

# Wave 2

data$treat2vC<-NA
data$treat2vC[data$treatment=="Control"]<-"Control"
data$treat2vC[data$treatment=="2: Baseline"]<-"Treatment 2"
m2vC<-lm(accessed_w2_c~treat2vC,data)
coeftest(m2vC,vcov=vcovHC(m2vC,type="HC2"))["treat2vCTreatment 2",]
confint.HC(m2vC,level=.95,parm="treat2vCTreatment 2",thevcov=vcovHC(m2vC,type="HC2"))


data$treat5vC<-NA
data$treat5vC[data$treatment=="Control"]<-"Control"
data$treat5vC[data$treatment=="5"]<-"Treatment 5"
m5vC<-lm(accessed_w2_c~treat5vC,data)
coeftest(m5vC,vcov=vcovHC(m5vC,type="HC2"))["treat5vCTreatment 5",]
confint.HC(m5vC,level=.95,parm="treat5vCTreatment 5",thevcov=vcovHC(m5vC,type="HC2"))

data$treat6vC<-NA
data$treat6vC[data$treatment=="Control"]<-"Control"
data$treat6vC[data$treatment=="6"]<-"Treatment 6"
m6vC<-lm(accessed_w2_c~treat6vC,data)
coeftest(m6vC,vcov=vcovHC(m6vC,type="HC2"))["treat6vCTreatment 6",]
confint.HC(m6vC,level=.95,parm="treat6vCTreatment 6",thevcov=vcovHC(m6vC,type="HC2"))

# Wave 3

data$treat8vC<-NA
data$treat8vC[data$treatment=="Control"]<-"Control"
data$treat8vC[data$treatment=="8"]<-"Treatment 8"
m8vC<-lm(accessed_w3_c~treat8vC,data)
coeftest(m8vC,vcov=vcovHC(m8vC,type="HC2"))["treat8vCTreatment 8",]
confint.HC(m8vC,level=.95,parm="treat8vCTreatment 8",thevcov=vcovHC(m8vC,type="HC2"))

data$treat10vC<-NA
data$treat10vC[data$treatment=="Control"]<-"Control"
data$treat10vC[data$treatment=="10"]<-"Treatment 10"
m10vC<-lm(accessed_w3~treat10vC,data)
coeftest(m10vC,vcov=vcovHC(m10vC,type="HC2"))["treat10vCTreatment 10",]
confint.HC(m10vC,level=.95,parm="treat10vCTreatment 10",thevcov=vcovHC(m10vC,type="HC2"))

# Wave 4
data$treat3vC<-NA
data$treat3vC[data$treatment=="Control"]<-"Control"
data$treat3vC[data$treatment=="3"]<-"Treatment 3"
m3vC<-lm(accessed_w4~treat3vC,data)
coeftest(m3vC,vcov=vcovHC(m3vC,type="HC2"))["treat3vCTreatment 3",]
confint.HC(m3vC,level=.95,parm="treat3vCTreatment 3",thevcov=vcovHC(m3vC,type="HC2"))

data$treat4vC<-NA
data$treat4vC[data$treatment=="Control"]<-"Control"
data$treat4vC[data$treatment=="4"]<-"Treatment 4"
m4vC<-lm(accessed_w4~treat4vC,data)
coeftest(m4vC,vcov=vcovHC(m4vC,type="HC2"))["treat4vCTreatment 4",]
confint.HC(m4vC,level=.95,parm="treat4vCTreatment 4",thevcov=vcovHC(m4vC,type="HC2"))
    



```


# Conclusions

Clear action steps and behaviorally informed messaging can be a significant help for military retirees reactivating myPay accounts. The best performing language will be utilized in DFAS’ future communications to further improve their service to military retirees.

# Appendix

Within any 6 day window the proportion accessing myPay in the control group (no initial email) is quite small (between 0.57 and 0.88 percent) compared to treatmement effects that range between a low of 3.03 percent (Group 7: Past Hassles) and a high of 6.33 percent (Increase in Security)

```{r}
with(data,apply(table(accessed_w1_c,treatment),2,prop.table))
with(data,apply(table(accessed_w2_c,treatment),2,prop.table))
with(data,apply(table(accessed_w3_c,treatment),2,prop.table))
with(data,apply(table(accessed_w4,treatment),2,prop.table))

```

## The effects of adding a signature:

```{r}
# Adding a signature has no effect relative to baseline email
m4<-lm(accessed_w4_34v2~treatment,data)
coeftest(m4,vcov=vcovHC(m4,type="HC2"))
```

## The effects of discussing tax benefits


```{r}
m5<-lm(accessed_w2_567v2~treatment,data)
coeftest(m5,vcov=vcovHC(m5,type="HC2"))
# None of the tax benefits appear more successful than the baseline email
```

## The effect of other benefits benefits

```{r}
# All of the general benefits appear more effecive than the baseline email
m6<-lm(accessed_w3_8910v2~treatment,data)
coeftest(m6,vcov=vcovHC(m6,type="HC2"))

```


