---
title: "myPay Analysis"
author: "Paul Testa"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r init,echo=F}
library(knitr)
opts_chunk$set(eval=T,echo=T,message=F,warning=F,cache=T)

```



# Agency Objective

Increase military retirees’ myPay reactivations before tax season in order to offset burden during tax season.

# Background

myPay is an online platform designed to help military retirees manage their military retirement finances. myPay allows convenient electronic download of tax documents, such as a Retirement Account Summary and IRS-1099 forms. While a myPay account is created for all military retirees automatically, many have never used it, or have let their accounts deactivate by not logging in for 180 consecutive days; about 55% of the retiree population have inactive myPay accounts (1.4 out of 2.5 million retirees). This poses a problem during tax season, when the number of retirees contacting the Defense Finance and Accounting Service (DFAS) surges as retirees try to either re-gain access to myPay or request additional copies of their tax documents (paper tax documents are mailed to retirees as default). This not only places substantial costs and burden on DFAS, but it also may result in slower services for the military retirees.

# Design

In collaboration with DFAS, the Social and Behavioral Sciences Team (SBST) designed nine different emails to encourage reactivation of myPay accounts. One email was designated as the baseline template, created based on existing language about myPay, with added clear action steps on how to reactivate an account. Eight email variations were built upon the baseline template using behavioral insights: an emails with the team’s signature versus an email with the director’s signature, three emails with different framings of benefits of using myPay during tax season, one with increased funds for our troops by cutting operation costs, one appealing to the security of using myPay, and one that provides social proof of retirees already using myPay. As a control, one group did not initially receive any emails.

Due to operational limitations, the emails were sent out in four separate waves between September 8 and November 5, 2015, two to three emails at a time. Emails were mostly sent on Tuesdays, with the exception of the final two emails, which went out on a Wednesday and the subsequent Thursday.
After the pilot, two emails were chosen from the original nine, and sent to the control group. Including the control group, this pilot included 261,754 retirees.


**Table 1. Treatment Conditions in Retiree myPay Account Reactivation Pilot**

| Treatment Group | Treatment                       | Date Sent | N     |
|---------------- |:-------------------------------:|-----------|------:|
| Group 1         | Control: No initial email$^a$       | 8-Dec-15$^a$  | 49496 |
| Group 2         | Baseline Template               | 22-Sep-15 | 23637 |
| Group 3         | Signature: Individual           | 4-Nov-15  | 23593 |
| Group 4         | Signature: Group                | 5-Nov-15  | 23538 |
| Group 5         | Tax Time Benefits: General      | 22-Sep-15 | 23551 |
| Group 6         | Tax Time Benefits: Specific     | 22-Sep-15 | 23765 |
| Group 7         | Tax Time Benefits: Past Hassles | 9-Sep-15  | 23580 |
| Group 8         | Benefits to Troops              | 20-Oct-15 | 23700 |
| Group 9         | Increase in Security            | 20-Oct-15 | 23455 |
| Group 10        | Social Proof                    | 20-Oct-15 | 23439 |

$^a$ After the initial treatment emails were sent, subject in the control condition were randomly assigned to receive the treatment of Group 6 (Tax Time Benefits: Specific), Group 9 (Increase in Security) or a combination of the messages in Group 6 and 9.

Subjects in a  list of 212,258 email addresses were completely randomly assigned to 1 of 10 conditions (9 treatment and 1 control).  Emails were sent  in waves.   Each wave contains two to three treatment conditions. Specifically:

**Table 2. Treatment Assignment by Waves**

| Wave | Date             | Treatments        | N     |
|------:|------------------:|-------------------:|-------:|
| 1    | 9/9/15           | Group 7           | 23580 |
| 2    | 9/22/15          | Groups 2, 5, & 6  | 70953 |
| 3    | 10/20/15         | Groups 8, 9, & 10 | 70594 |
| 4    | 11/4/15; 11/5/15 | Groups 3 & 4 | 47131 |

# Analysis Plan

Here is the plan for the analysis. We agreed on this plan before doing analysis.


## Data and Outcomes

Six to nine days after each treatment wave was sent, DFAS queried their myPay records  providing an updated list of who had accessed and reactivated their myPay accounts.  The outcome for a particular treatment is defined from these multiple datasets as the proportion of recipients who accessed their myPay account over a six-day period starting on the day a given email treatment was sent. This outcome can be compared to  the proportion of participants accessing their myPay accounts  over the same number of days (but at a potentially different point in time)  in the following groups:

- All untreated participants (Participants who have not yet received an email)
- All “control” participants (Group 1)  who did not receive an email until December 8, 2015.
- Subjects in the baseline treatment (Group 2)
- Subjects in a similar treatment condition  (e.g. Group 3 vs Group 4)

Table 4 details the dates data were collected, the treatments administered at that time  and  the relevant comparisons for each treatment. Note the date ranges for some comparisons involve comparisons across different waves of the study.

**Table 4: Dates of treatment administration and outcome observation**

| Wave | Date                    | Treatment             | Relevant Comparisons           |
|------|-------------------------|-----------------------|--------------------------------|
|      |                         |                       |                                |
| 1    | Treatment Administered: | Group 7               | All Untreated                  |
|      | 9/9/15                  |                       | Groups 1-6, 8-10               |
|      |                         | Outcome defined over: | Dates: 9/9-9/14                |
|      | Outcomes Observed: 9/18 | 9/9-9/14              | Wave: 1                        |
|      | 9/18/15                 |                       | Control                        |
|      |                         |                       | Group 1                        |
|      |                         |                       | Dates: 9/9-9/14                |
|      |                         |                       | Wave: 1                        |
|      |                         |                       | Baseline                       |
|      |                         |                       | Groups 2                       |
|      |                         |                       | Dates: 9/22-9/27               |
|      |                         |                       | Wave: 2                        |
|      |                         |                       | Similar Treatments             |
|      |                         |                       | Groups 5, 6                    |
|      |                         |                       | Dates: 9/22-9/27               |
|      |                         |                       | Wave: 2                        |
|      |                         |                       |                                |
| 2    | Treatment Administered: | Groups 2, 5, & 6      | All Untreated                  |
|      | 9/22/15                 |                       | Groups 1,3,4, 8-10             |
|      |                         | Outcome defined over: | Dates: 9/22-9/27               |
|      | Outcomes Observed:      | 9/22-9/27             | Wave: 2                        |
|      | 9/28/15                 |                       | Control                        |
|      |                         |                       | Group 1                        |
|      |                         |                       | Dates: 9/22-9/27               |
|      |                         |                       | Wave: 2                        |
|      |                         |                       | Baseline                       |
|      |                         |                       | Groups 2                       |
|      |                         |                       | Dates: 9/22-9/27               |
|      |                         |                       | Wave: 2                        |
|      |                         |                       | Similar Treatments             |
|      |                         |                       | Groups 5, 6,7                  |
|      |                         |                       | Dates: 9/22-9/27               |
|      |                         |                       | Wave: 2                        |
| 3    | Treatment Administered: | Groups 8, 9, & 10     | All Untreated                  |
|      | 10/20/15                |                       | Groups 1,3,4                   |
|      |                         | Outcome defined over: | Dates: 10/20-10/25             |
|      | Outcomes Observed:      | 10/20-9/25            | Wave: 3                        |
|      | 10/28/15                |                       | Control                        |
|      |                         |                       | Group 1                        |
|      |                         |                       | Dates: 10/20-10/25             |
|      |                         |                       | Wave: 3                        |
|      |                         |                       | Baseline                       |
|      |                         |                       | Groups 2                       |
|      |                         |                       | Dates: 9/22-9/27               |
|      |                         |                       | Wave: 2                        |
|      |                         |                       | Similar Treatments             |
|      |                         |                       | Groups 8-10                    |
|      |                         |                       | Dates: 10/20-10/25             |
|      |                         |                       | Wave: 3                        |
| 4    | Treatment Administered: | Groups 3              | All Untreated                  |
|      | 11/4/15;                |                       | Groups 1                       |
|      | 11/5/15                 | Outcome defined over: | Dates: 11/4-11/9 [11/5-11/10]  |
|      |                         | 11/4-11/9             | Wave: 4                        |
|      | Outcomes Observed:      |                       | Control                        |
|      | 11/10/15                | Group 4               | Group 1                        |
|      |                         | Outcome defined over: | Dates: 11/4-11/10 [11/5-11/10] |
|      |                         | 11/5-11/10            | Wave: 4                        |
|      |                         |                       | Baseline                       |
|      |                         |                       | Groups 2                       |
|      |                         |                       | Dates: 9/22-9/27               |
|      |                         |                       | Wave: 2                        |
|      |                         |                       | Similar Treatments             |
|      |                         |                       | Groups 3,4                     |
|      |                         |                       | Dates: 11/4-11/10 [11/5-11/10] |
|      |                         |                       | Wave: 4                        |
| 5    | Treatment Administered: | Group 1               | All Untreated                  |
|      | 12/8/15;                |                       | Group 1                        |
|      |                         | Outcome defined over: | Dates: 11/4-11/9 [11/5-11/10]  |
|      |                         | 12/8-12/13            | Wave: 4                        |
|      | Outcomes Observed:      |                       | Control                        |
|      | 12/4/15                 |                       | Group 1                        |
|      |                         |                       | Dates: 11/4-11/10 [11/5-11/10] |
|      |                         |                       | Wave: 4                        |
|      |                         |                       | Baseline                       |
|      |                         |                       | Group 2                        |
|      |                         |                       | Dates: 9/22-9/27               |
|      |                         |                       | Wave: 2                        |
|      |                         |                       | Similar Treatments             |
|      |                         |                       | Groups 6,9                     |
|      |                         |                       | Dates: 9/22-9/27; 10/20-10/25  |
|      |                         |                       | Wave: 2;3                      |


The aim is to compare the relative effectiveness of 9 treatments in terms of encouraging retirees to access their myPay accounts within seven days of receipt of the email encouragement.

We will report the average proportion subscribing in each of the 9 treatment conditions compared to those in the relevant comparison groups.

The first question is whether each treatments differs from the control group, adjusting for multiple comparisons .

Then we will test the following hypotheses relative to the baseline template, again adjusting for multiple comparisons:

The effects of adding a signature:

- 2 versus 3 and 4 (Baseline vs Any Signature)
- 2 versus 3  (Baseline vs Specific Signature)
- 2 versus 4  (Baseline vs General Signature)
- 3 versus 4  (Specific vs General Signature)

The effect of tax benefits

- 2 versus 5,6,7 (Baseline vs Tax Benefits)
- 2 versus 5  (Baseline vs General Benefits)
- 2 versus 6  (Baseline vs Specific Benefits)
- 2 versus 7  (Baseline vs Past Hassles)
- 5 versus 6  (General Benefits vs Specific Benefits)
- 5 versus 7  (General Benefits vs Past Hassles)
- 6 versus 7  (Specific Benefits vs Past Hassles)

The effect of other benefits benefits

- 2 versus 8,9,10 (Baseline vs Other Benefits)
- 2 versus 8  (Baseline vs Troop Benefits)
- 2 versus 9  (Baseline vs Security Benefits)
- 2 versus 10  (Baseline vs Social Proof)
- 8 versus 9  (Troop Benefits vs Security Benefits)
- 8 versus 10  (Troop Benefits vs Past Hassles)
- 9 versus 10  (Security Benefits vs Social Proof)

We do not have expected directions of effects, so we use two-tailed p-values throughout.

Finally, we will also explore the potential for heterogenous effects within branch of service and age group.


# Results

```{r}
data<-read.csv("data/mypay_clean.csv")
# Relevel treatment factors
data$treatment<-factor(data$treatment,levels=c("Control","2: Baseline",3:10))

```


Between September 8 and November 10, 2015, 5.5% of retirees that did not receive any emails reactivated their myPay accounts.

```{r}
with(data,apply(table(accessed,treat_anyemail),2,prop.table))

```



In comparison, 10.2% of retirees that received a treatment email had reactivated their myPay accounts – a difference of 4.7 percentage points from the control group (p < 0.001, 95% CI [4.42,4.99]).

```{r}
m1<-lm(accessed~treat_anyemail,data)
summary(m1)
with(data,prop.test(table(treat_anyemail,accessed)[2:1,2:1]))
# Slightly different confidence intervals, same result
```

The most effective email was the one that appealed to the security of using myPay, with 11.1% reactivation rate – an increase of 5.6 percentage points (p < 0.001, 95% CI [5.16,6.06])).

```{r}
with(data,apply(table(accessed,treatment),2,prop.table))[,9]
library(lmtest)
m2<-lm(accessed~treatment,data)
coeftest(m2)["treatment9",]
confint(m2)["treatment9",]

data$treat9vC<-NA
data$treat9vC[data$treatmen=="Control"]<-"Control"
data$treat9vC[data$treatmen=="9"]<-"9"
with(data,
     prop.test(table(treat9vC,accessed)[1:2,2:1]))
```


Interestingly, the email that used specific tax document names (i.e. RAS and 1099) performed slightly better than the email that used the generic phrase “tax documents” by 0.6 percentage points (p < 0.05, 95% CI [0.10, 1.18]) .

```{r}
data$treat_5v6<-NA
data$treat_5v6[data$treatment=="5"]<-"General"
data$treat_5v6[data$treatment=="6"]<-"Specific"
m3<-lm(accessed~treat_5v6,data)
summary(m3)
confint(m3)
with(data,
     prop.test(table(treat_5v6,accessed)[1:2,2:1]))

```

We also estimated that the emails led to 800-1,000 additional customer service calls during the 3.5 months of pilot (out of 261,754 retirees). A follow-up analysis post-tax season will allow us to determine whether the increase in call volumes during the pilot will have offset the burden during tax season.


# Conclusions

Clear action steps and behaviorally informed messaging can be a significant help for military retirees reactivating myPay accounts. The best performing language will be utilized in DFAS’ future communications to further improve their service to military retirees.
Follow-up analysis will allow us to determine whether this pilot had impact on DFAS’ costs and burden during tax season.

# Appendix

Within any 6 day window the proportion accessing myPay in the control group (no initial email) is quite small (between 0.57 and 0.88 percent) compared to treatmement effects that range between a low of 3.03 percent (Group 7: Past Hassles) and a high of 6.33 percent (Increase in Security)

```{r}
with(data,apply(table(accessed_w1_c,treatment),2,prop.table))
with(data,apply(table(accessed_w2_c,treatment),2,prop.table))
with(data,apply(table(accessed_w3_c,treatment),2,prop.table))
with(data,apply(table(accessed_w4,treatment),2,prop.table))

```

## The effects of adding a signature:

```{r}
# Adding a signature has no effect relative to baseline email
summary(lm(accessed_w4_34v2~treatment,data))
```

## The effects of discussing tax benefits


```{r}

summary(lm(accessed_w2_567v2~treatment,data))
# None of the tax benefits appear more successful than the baseline email
```

## The effect of other benefits benefits

```{r}
# All of the general benefits appear more effecive than the baseline email
summary(lm(accessed_w3_8910v2~treatment,data))

```


